---
title: "HW05 Part 2: Importing"
author: "Christina Beauford"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```



##Limpets
```{r}
limpets <- read_csv("C:/Users/Christina/Documents/BI485/beauford_christina/hw05/limpets.csv")
```

```{r}
limpets
```

No, the data is not tidy. There are multiple of the same rows of the same lengths recorded. The rows should be numbered or there should be a number where it lists how many of the same entry there are to ensure that this data is supposed to be multiplied like this. 

```{r}
read_csv("limpets.csv", col_types = c("ncc"))
```

```{r}
sites_col <- limpets$Site <- factor(limpets$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO",  "CTZ1", "CTZ2", "CNM"))
```

```{r}
ggplot(data = limpets, mapping = aes(x = Site, y = Length, color = Protected)) +
  geom_boxplot()
```

```{r}
ggplot(data = limpets, mapping = aes(x = Site, y = Length, color = Protected)) +
  geom_boxplot() +
  labs(x = "Site", y = "Length (mm)")
```
The sites WP and DP have outliers. The protected limpets seem to be larger in length.

##Roseate Terns

```{r}
roseate_terns <- read.delim("C:/Users/Christina/Documents/BI485/beauford_christina/hw05/roseate_terns.txt")
```

```{r}
roseate_terns
```

The data looks tidy, except there are some NA's listed throughout the data.

```{r}
tidy_roseate <- roseate_terns %>% 
  gather("AR", "AC", "BR", "BL", "BC", "JR", "JL", "JC", key = "Site", value = "Individuals")

tidy_roseate
```

```{r}
filter(tidy_roseate, is.na(Individuals) == FALSE)
```

```{r}
ggplot(data = tidy_roseate, mapping = aes(x = year, y = Individuals)) +
  geom_line() +
  facet_wrap(~ Site, nrow = 2)
```

```{r}
ggplot(data = tidy_roseate, mapping = aes(x = year, y = Individuals)) +
  geom_line() +
  facet_wrap(~ Site, nrow = 2) +
  labs(x = "Year", y = "Roseate Terns")
```

AR's population significantly increased between 1992 and 2016. AC, BC, BL, JC, and JL slightly increased the population between 1992 and 2016.JR and BR decreased their populations during this time.

The data in the ones with breaks probably say NA or the data is inconclusive for those years where the breaks are located.

##Darters

```{r}
darters <- read.delim("C:/Users/Christina/Documents/BI485/beauford_christina/hw05/darters.txt")
```


```{r}
darters
```



















##Blacklip Abalone
```{r}
abalone <- read_csv("C:/Users/Christina/Documents/BI485/beauford_christina/hw05/abalone.csv")
```


```{r}
abalone
```

```{r}
abalone1 <- select(abalone, 2:10)
abalone1
```

```{r}
ggplot(data = abalone1, mapping = aes(x = Type, y = Height)) +
  geom_boxplot()
```

```{r}
ggplot(data = abalone1, mapping = aes(x = Rings, y = Height)) +
  geom_point()
```
No, the large outliers are not extremely old.


```{r}
abalone2 <- filter(abalone1, Height < 0.3)

abalone2
```

```{r}
abalone_types <- abalone2$Type <- factor(abalone2$Type, levels = c("I", "F", "M"), ordered = TRUE)
```

```{r}
ggplot(data = abalone2, mapping = aes(x = Rings, y = Height, color = Type)) +
  geom_point()
```

Generally, With increasing rings, there is an increase in height. When the ring number is just before 20, the height decreases slightly. Males and females seem to tie with largest height and immatures are definitely smaller in height, but not all females and males are larger than immatures.

```{r}
filter(abalone2, Height == 0)
```

```{r}
abalone2[abalone2 == 0] <- NA
```

```{r}
abalone3 <- filter(abalone2, is.na(Height) == FALSE)
```

```{r}
abalone3
```

```{r}
ggplot(data = abalone3, mapping = aes(x = Rings, y = Height, color = Type)) +
  geom_point()
```

```{r}
ggplot(data = abalone3, mapping = aes(x = Diameter, y = Height, color = Type)) +
  geom_point()
```

```{r}
ggplot(data = abalone3, mapping = aes(x = WholeWeight, y = ShuckedWeight, color = Type)) +
  geom_point()
```

