---
title: "HW 06 Part 2: Wrangling II"
author: "Christina Beauford"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(knitr)
library(smatr)
```

##Darters in Riffles

```{r}
darters <- read_csv("C:/Users/Christina/Documents/BI485/beauford_christina/hw06/darters.txt")
```

```{r}
darters
```
```{r}
start_pos <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
stop_pos <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 42)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
darters1 <- read_fwf("darters.txt", fwf_positions(start_pos, stop_pos, col_names = names), skip = 15)
```

```{r}
darters1
```

```{r}
darters2 <- darters1 %>%
  filter(!species %in% c("zonale", "tetrazonum"))
darters2
```

```{r}
mutate(darters2,
  total_substrate = majsub + minsub)
```
Yes, the total substrate sums to 1 in all of the observations.


```{r}
mutate(darters2,
       riffle = ifelse(riffle == 1, 
                       "Riffle 1", 
                       "Riffle 2")
       )
```

```{r}
mutate(darters2,
       length = length * 10)
```

```{r}
mutate(darters2,
       sex = ifelse(sex == "f",
                           "Female",
                           "Male"))
```

```{r}
darters3 <- group_by(darters2, species)
darters_summary <- darters3 %>%
  summarise(length = mean(length, na.rm = TRUE),
          depth = mean(depth, na.rm = TRUE),
          velocity = mean(velocity, na.rm = TRUE))
darters_summary
```


Plot 1
```{r}
ggplot(data = darters3, mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median) +
  geom_point(data = darters_summary, mapping = aes(x = species, y = length), shape = 15)  +
  coord_flip() +
  facet_wrap(~ riffle, nrow = 2) +
  labs(x = "Species", y = "Length (mm)")
```

In riffle 1 compared to riffle 2, spectabile's range seemed to stay close to the same, but in riffle 2, the length is slightly smaller. In riffle 2, caeruleum has a slightly smaller range and the length is slightly smaller in general. In riffle 2, blennioides range is significantly larger and the length is much larger in general too.


Plot 2
```{r}
ggplot(data = darters3, mapping = aes(x = species, y = length)) +
  stat_summary(
    fun.y = median,
    fun.ymin = min,
    fun.ymax = max,
    geom = "point",
    size = 2.2) +
  stat_summary(
    fun.y = mean,
    fun.ymin = min,
    fun.ymax = max,
    size = 3,
    color = "maroon") +
  geom_point(data = darters_summary, mapping = aes(x = species, y = length), shape = 15, color = "gray70") +
  facet_wrap(~ riffle, nrow = 2) +
  labs(x = "Species", y = "Length (mm)") +
  coord_flip()
```

##Stress and Corticosterones in Birds

```{r}
library(tidyverse)
library(smatr)
```

```{r}
names <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")
feathers <- read_csv(("C:/Users/Christina/Documents/BI485/beauford_christina/hw06/feathers.csv"), locale = locale(decimal_mark = ","), skip = 1, col_names = names) 

```

The data has two different dates listed, which seems weird. It does not seem to be tidy.

```{r}
feathers
```

```{r}
mean_tarsus <- mean(feathers$tarsus)
mean_tarsus
```

```{r}
log_mass <- log(feathers$mass)
log_tarsus <- log(feathers$tarsus)

log_mass
log_tarsus

mutate(feathers, log_mass, log_tarsus)
feathers
```

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = feathers)
major_axis
```
```{r}
summary(major_axis)
```

```{r}
coef(major_axis)
```

```{r}
b <- 1.864889
```

```{r}
new_smi <- (feathers$mass * ((mean_tarsus/feathers$tarsus)^b))
new_smi
```
```{r}
feathers$smi
```
The values do not differ more than 0.1.

```{r}
mean_basal_cort <- mean(feathers$basal_cort)
mean_stress_cort <- mean(feathers$stress_cort)
sd_basal_cort <- sd(feathers$basal_cort, na.rm = TRUE)
sd_stress_cort <- sd(feathers$stress_cort, na.rm = TRUE)
se_basal_cort <- sd_basal_cort/(sqrt(43))
se_stress_cort <- sd_stress_cort/(sqrt(43))

se_basal_cort
se_stress_cort
```

