---
title: "HW 06 Part 1: Complete the sections"
author: "Christina Beauford"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

```{r}
library(tidyverse)
library(smatr)
```

##Coding Style

```{r}
file_path <- "data/"
```



```{r}
data(iris)
ggplot(data = iris, mapping = aes(x = Petal.Length, y = Sepal.Length)) +
  geom_point(mapping = aes(color = Species, shape = Species)) +
  geom_smooth(mapping = aes(color = Species))
```

```{r}
ggplot(data = iris,
       aes(x = Petal.Length,
           y = Sepal.Length)) +
  geom_point(aes(color = Species,
                 shape = Species)) +
  geom_smooth(aes(color = Species))
```


##Wrangling II

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

```{r}
library(nycflights13)
library(tidyverse)
```

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 

Do not answer any of the questions.

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
```

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```
mutate() adds new columns that are functions of existing columns. It adds the columns at the end of the dataset.

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
Refers to columns that were just created with mutate().


```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
transmutate() keeps only the new columns created.


```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```
Showing that the mutate() variables must be vectorized, they can be used with arithmetic functions. %/% is integer division (used in this code chunk).


```{r}
(x <- 1:10)

lag(x)

lead(x)
```
lead and lag allow you to refer to leading and lagging values. Most useful when used with group_by().


```{r}
x

cumsum(x)

cummean(x)
```
"cumulative and rolling aggregates". cumsum, cummean are cumulative functions.


```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)

min_rank(desc(y))
```
Ranking. min_rank is the usual type of ranking (1st, 2nd, 3rd...); gives the smallest values first. desc gives largest values first (descending).


```{r}
row_number(y)

dense_rank(y)

percent_rank(y)

cume_dist(y)
```
row_number: ties.method = "first"
dense_rank: min_rank, but without gaps between ranks
percent_rank: rescales the min_rank to a number between 0 and 1
cume_dist: cumulative distrubution function; proportion of all values <= current rank


### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```
summarise() collapses a data frame to a single row.


```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```
average delay per date provided by group_by and summarise. This shows that summarise is more useful with group_by. This changes the reference to individual groups rather than the complete dataset. Using dplyr verbs on group_by automatically applies by group. 


```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")
```
relationship between distance and average delay for each location.


```{r}
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```
plot to visualize above codes.

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```
same problem as above, but using pipes.

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
This shows the many missing values "NA". 
```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```
This removes the values "NA" prior to computing.
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
This removes the missing values for the cancelled flights.

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
```

```{r}
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```
Highest average delays by the tail number of the plane.

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
```

```{r}
ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```
Scatterplot of number of flights vs average delay.

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```
Shows less of the extreme variation in small groups.


### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```
Shows how filter() can be useful with group_by. Finding the worse members of each group.


```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```
All groups bigger than a threshold; using filter and group_by.


```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```
Standardise to compute per group metrics; uses filter(), mutate(), and select().

